Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-19 doc/GMP2015/appendix_a.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

    \item 4 cases: \todo{Uses 2D requirement:} $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.

@@

    \item 1 case: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible. Since $C_i\cup C_k\neq C$ (and $C_j\cup C_k\neq C$), we know by the same logic as the previous case that the $C_i\cap C_k$ must be trivial (a single node). However, $C_i\cap C_k=\inducedOnC{R'_j\cup D'_{i,j}}$. This causes a contradiction, the intersection cannot be trivial because $R'_j$ and $D'_{i,j}$ are not empty sets and are disjoint.

    \item 2 cases: \todo{Uses 2D requirement:} $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible. The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.

>>>>

    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.

@@

    \item 1 case: \usestwod $C_k=\inducedOnC{R'_i\cup R'_j\cup D'_{i,j}}$ is not possible. Since $C_i\cup C_k\neq C$ (and $C_j\cup C_k\neq C$), we know by the same logic as the previous case that the $C_i\cap C_k$ must be trivial (a single node). However, $C_i\cap C_k=\inducedOnC{R'_j\cup D'_{i,j}}$. This causes a contradiction, the intersection cannot be trivial because $R'_j$ and $D'_{i,j}$ are not empty sets and are disjoint.

    \item 2 cases: \usestwod $C_k=\inducedOnC{R'_i\cup R_j\cup D'_{i,j}}$ and $C_k=\inducedOnC{R_i\cup R'_j\cup D'_{i,j}}$ are not possible. The proof mirrors the previous case, except here you must choose $C_i$ and $C_j$ respectively.

### 2014-12-19 doc/GMP2015/s01_introduction.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system. Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems \uncited. When the required
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in \uncited\uncited (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems \uncited)
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems).

>>>>

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system
\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
  Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems
\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
  and to find paths within the connected components. When the
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in
\cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII}
  (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems
  \cite{sitharam2014incidence})
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems) \uncited.

### 2014-12-19 doc/GMP2015/s02_preliminaries.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

\subsection{Decomposition-Recombination (DR) Plans}
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%    \item \textbf{over-constrained} if it is not independent.
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
% \begin{definition}
%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
% \begin{definition}
%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}), is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
        set of $C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
%an isostatic vertex-maximal proper subgraph.
    %Note that nodes will be referred to interchangeably as ``the node that
    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
    For a given graph, there could be exponentially many  DR-plans and
    even optimal DR-plans in the size of the graph . A complete DR-plan
    is unique but may not be (is usually not) optimal. DR-plans
    of self-similar graphs are self-similar. More than one node (leaf)
    in a DR-plan forest may represent the same subgraph (vertex) of
    $G$.
\end{remark}

>>>>

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{over-constrained} if it is not independent.
%
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
        set of $C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
        We denote a complete DR-plan of $G$ as {\it
        ComDRP}$(G)$
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
    We denote an optimal DR-plan of $G$ as {\it OptDRP}$(G)$.

%
%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
%an isostatic vertex-maximal proper subgraph.
    %Note that nodes will be referred to interchangeably as ``the node that
    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}
%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
    For a given graph, there could be exponentially many  DR-plans and
    even optimal DR-plans in the size of the graph. A complete DR-plan
    is unique but may not be (is usually not) optimal. DR-plans
    of self-similar graphs are self-similar. More than one node (leaf)
    in a DR-plan forest may represent the same subgraph (vertex) of
    $G$.
\end{remark}

### 2014-12-19 doc/GMP2015/s03_drp.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.

\section{The canonical DR-plan}
\label{sec:DRP}

\subsection{Theory}

% In the problem of the optimal DR-Plan there is generally not a unique plan.
% % Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same. Therefore, all choices of decomposition are in some sense equivalent. The theorem we seek to prove is thus:
% However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems. Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
% % To assist in showing this, we prove this core theorem throughout this section:

@@

In the problem of the optimal DR-Plan there is generally not a unique plan. Furthermore, there is no obvious way to find the optimal plan. In this section we present a particular way to construct a plan that guarantees optimality. We call this the \dfn{canonical} plan.

@@

The importance of this formulation is stated succinctly in the following theorem.

\begin{theorem}\label{theorem:canonical_is_optimal}
    A canonical DR-plan is an optimal DR-plan.
\end{theorem}

The proof of this theorem is a direct result of this more general theorem.

\begin{theorem}\label{theorem:main}
Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$ with children $C_1,\ldots,C_N$ keep children according to the following rules.
\begin{enumerate}
    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
\end{enumerate}
This covers all cases and the resulting plan is an optimal DR-plan.
\end{theorem}

\begin{remark}
    \todo{this captures cluster minimality}
\end{remark}

This theorem (proven in the Appendix) relies on the following crucial observations and lemmas.

\begin{observation}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
\end{observation}

\begin{lemma}\label{lemma:wc_intersection_is_C}
$C_i\cup C_j$ is well-constrained if and only if $C_i\cup C_j = C$.
\end{lemma}

\begin{lemma}\label{lemma:wc_intersection_makes_all_wc}
If $C_i\cup C_j$ is well-constrained, then $\forall k: C_i\cup C_k$ is well-constrained.
Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
\end{lemma}

\begin{lemma}\label{lemma:uc_intersection_makes_all_uc}
If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{lemma}

>>>>

\section{The canonical DR-plan}
\label{sec:DRP}
% In the problem of the optimal DR-Plan there is generally not a unique plan.
% % Indeed, we will prove a union of $N$ well-constrained subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same. Therefore, all choices of decomposition are in some sense equivalent. The theorem we seek to prove is thus:
% However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems. Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
% % To assist in showing this, we prove this core theorem throughout this section:

@@

The NP-hardness of the DR-planning problem for 2D bar-joint graphs is partly
the consequence of possibly exponential number of DR-plans. On
the other hand, although the complete DR-plan is unique
it could have large average fan-in and exponentially many nodes making it
far from optimal.

In this section we define a \dfn{canonical} plan to capture those aspects of an optimal
DR-plan that mimic a complete DR-plan's uniqueness, while for independent
(underconstrained or isostatic graphs), we show that the nonunique parts do not affect
optimality.

@@

This definition gives a canonical DR-plan  a surprisingly strong Church-Rosser property.

\begin{theorem} \label{theorem:canonical_is_optimal}
    \label{theorem:main}
    Any canonical DR-plan is an optimal DR-plan.
\end{theorem}

%The proof of this theorem is a direct consequence of the following  more general theorem.

%\begin{theorem}\label{theorem:main}
%Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$
%with children $C_1,\ldots,C_N$ preserve children according to the following rules.
%\begin{enumerate}
%    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
%    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
%\end{enumerate}
%The resulting canonical DR-plan is an optimal DR-plan.
%\end{theorem}

This theorem (proven in the Appendix) relies on the following crucial observation and lemma
that are also used for other types of qusecs in the Application sections of
the paper.

\begin{observation}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are isostatic subgraphs of the same
isostatic graph $F$, \
then the following hold:
(1) $F_i\cup F_j$ is not trivial;
(2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial;
(3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and
(4) $F_i\cap F_j$ is not under-constrained.
\end{observation}

Then the following key properties hold at the nodes of a canonical DR-plan.

\begin{lemma}
Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
Then
\begin{enumerate}
    \item
    \label{lemma:wc_intersection_is_C}
$C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.
\item
    \label{lemma:wc_intersection_makes_all_wc}
If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic
Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
\item
    \label{lemma:uc_intersection_makes_all_uc}
If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{enumerate}
\end{lemma}

\begin{remark}
The first item in the above lemma generalizes to the union of any number of
children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{Cluster
Minimality},
defined in \cite{hoffman2001decompositionI}
holding for these canonical-optimal DR-plans.
\end{remark}
